---
title: "Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(haven)

```


##Read in some data

Read in the litters dataset

```{r read_in_litters}
litters_df = read_csv("./data/FAS_litters.csv")
#read_csv is a fucntion in the readr package
#csv=comma, seperated values. 
#the "." means from the folder where the R project file is located, aka. start here
litters_df

#below start the data cleaning from rename the variables:
litters_df = janitor::clean_names(litters_df)
#janitor is another package that need to be loaded here, double ":" to load, and that use the function "clean_name" in jamitor to rename the variables into snake_form. 

```

## Take a look at the data

Print in console
```{r}
litters_df
```

```{r}
head(litters_df)
tail(litters_df)
```
```{r}
skimr::skim(litters_df)
#give a brief look of the basic data descriptive infomation like missing, means, and histgram, ect. 
#also wa can type "view(litters_df"in cosole instead of rmd chunk to open a new window of the dataset. (open new window in rmd will make knit harder)
```



## Options in read.csv

One can always check out the help tab (?read.csv()) for more details by typing ?read_csv in console

```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE)
#skip the first 10 rows (including the variable name row), can use this if the start rows have some mess
litters_df


litters_df = read_csv("./data/FAS_litters.csv", na = c("", "NA", ".", 999))
#for all items in the collection (c means collection), R will treat that as missing data. 
```


##read in other format
###read in excel file
```{r}
library(readxl)#also in the top of the file, integrate all the package that will be used more than once
mlb_df = read_excel("./data/mlb11.xlsx")
mlb_df

#use ?read_excel() to find the specific function like range and sheet for excel file
mlb_df = read_excel("./data/mlb11.xlsx", range = "A1:F7")
```


###read in SAS file
```{r}
library(haven)
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
```


##comparison with base R

what about read.csv...? which is build in R and no need to load packages.
well, neven use that. 

```{r}
litters_base = read.csv("./data/FAS_litters.csv")
litters_readr = read_csv("./data/FAS_litters.csv")

litters_base
litters_readr


```

##Export data

Like, export thr mlb sub-table
```{r}
write_csv(mlb_df, "./data/mlb_subtable.csv")
```

